---
title: "Python Notebook for S-map for DE and ODE models"
output: html_notebook
---

### [0] Setting Python environment (in RStudio)
Download Python environment: https://www.python.org/downloads/

#### [0.1] Loading the package for Python environment
```{r}
library(reticulate)
```

#### [0.2] Installing libraries & Loading classes
In R chunk
```{r}
library_name = "pandas"
if (py_module_available(library_name)) {
  message(paste("The library", library_name, "is already installed."))
} else {
  message(paste("The library", library_name, "is not installed. Installing now..."))
  py_install(library_name)
}

library_name = "matplotlib"
if (py_module_available(library_name)) {
  message(paste("The library", library_name, "is already installed."))
} else {
  message(paste("The library", library_name, "is not installed. Installing now..."))
  py_install(library_name)
}

library_name = "pyEDM"
if (py_module_available(library_name)) {
  message(paste("The library", library_name, "is already installed."))
} else {
  message(paste("The library", library_name, "is not installed. Installing now..."))
  py_install(library_name)
}
```

In Python chunk
```{python}
import numpy as np   #for numerical calculations
import pandas as pd
import math    #for mathematical functions
import pyEDM   #for EDM analysis
import matplotlib.pyplot as plt   #for graphics
import multiprocessing as mp #for paralell computing
from numpy import exp
```

### [1] General functions for ODEs
4th-order explicit Runge-Kutta method with a fixed interval function

```{python}
def rk4(in_vec, out_vec, time, h_interval, dim, diff_vec, ref_data = None):
    #in_vec: the vector with the value at "time"
    #out_vec: the vector with which the updated vector value after one time step is saved
    #time: time value
    #h_interval: time step for discritizing ODE
    #dim: dimension of the ODE
    #diff_vec: function that determines the r.h.s. of the ODE

    k1 = np.zeros([dim])
    k2 = np.zeros([dim])
    k3 = np.zeros([dim])
    k4 = np.zeros([dim])

    temp_vec = np.zeros([dim])
    h_half = h_interval / 2.0
    t_h = time + h_half

    k1 = diff_vec(in_vec, time, h_interval, dim, ref_data)  # calculate k1
    temp_vec = in_vec + h_half * k1
    k2 = diff_vec(temp_vec, t_h, h_interval, dim, ref_data)  # calculate k2
    temp_vec = in_vec + h_half * k2
    k3 = diff_vec(temp_vec, t_h, h_interval, dim, ref_data)  # calculate k3
    temp_vec = in_vec + h_interval * k3
    k4 = diff_vec(temp_vec, time + h_interval, h_interval, dim, ref_data)  # calculate k4

    for i in range(0, dim):
      out_vec[i] = in_vec[i] + (h_interval / 6.0) * (k1[i] + 2.0 * k2[i] + 2.0 * k3[i] + k4[i])
```
### [2] Model01 setting
Interaction strength estimation for time-discrete model (DE) framework

#### [2.1] Model equations
2-Host 2-Parasitoid System with natural mortality (i.e., with overlapping generations)

Note: When the natural mortality goes to infinity, the model converges to 2-Host 2-Parasitoid Nicolson-Baily Model

\[
H_{k,t+1} = H_{k,t} + r_k exp[-a_{k}H_{k,t} - c_{1k} P_{1,t} - c_{2k} P_{2, t}] H_{k,t} - (1 - exp[-a_{k}H_{k,t} - c_{1k} P_{1,t} - c_{2k} P_{2,t}  - m_{H,k}]) H_{k,t}  \quad (k=1,2) \tag{D1}
\]
It is set as 
\[
H_{k,t+1} = H_{k,t} + F_{Hk}(H_{1,t}, H_{2,t}, P_{1,t}, P_{2,t})
\]

\[
P_{j,t+1} = P_{j,t} + \sum_{k=1,2} exp(-a_k H_{k,t}) (1 -  exp[-c_{1k} P_{1,t} - c_{2k} P_{2,t}]) \frac{c_{jk} P_{j,t}}{c_{1k} P_{1,t} + c_{2k} P_{2,t}} H_{k,t} -(1 - exp(-m_{P,j})) P_{j,t} \quad (j=1,2) \tag{D2}
\]
It is set as 
\[
P_{j,t+1} = P_{j,t} + F_{Pj}(H_{1,t}, H_{2,t}, P_{1,t}, P_{2,t})
\]


#### [2.2] Parameters and settings for the 4-species coupled host-parasitoid model

##### [2.2.1] Parameter value setting
```{python}
r1 = 2.0
r2 = 2.0
a1 = 0.1
a2 = 0.1
c11 = 0.3 #P1 -> H1
c21 = 0.1 #P2 -> H1
c12 = 0.1 #P1 -> H2
c22 = 0.3 #P2 -> H2
mH1 = 0.1
mH2 = 0.1
mP1 = 0.1
mP2 = 0.1
```
##### [2.2.2] Setting for numerical calculations
```{python}
j_Host1 = 0
j_Host2 = 1
j_Paras1 = 2
j_Paras2 = 3
```
##### [2.2.3] Functions specific to 4-sp model
Note that FH1, FH2, FP1, FP2, are defined as the changes of the abundance (X_t+1 - X_t)
```{python}
def FH1(in_vec, t):
  term1 = r1*exp(-a1 * in_vec[j_Host1] - c11 * in_vec[j_Paras1] - c21 * in_vec[j_Paras2]) * in_vec[j_Host1]
  term2 = (1.0 - exp(-a1 * in_vec[j_Host1] - c11 * in_vec[j_Paras1] - c21 * in_vec[j_Paras2] - mH1)) * in_vec[j_Host1]
  return term1 - term2

def FH2(in_vec, t):
  term1 = r2*exp(-a2 * in_vec[j_Host2] - c12 * in_vec[j_Paras1] - c22 * in_vec[j_Paras2]) * in_vec[j_Host2]
  term2 = (1.0 - exp(-a2 * in_vec[j_Host2] - c12 * in_vec[j_Paras1] - c22 * in_vec[j_Paras2] - mH2)) * in_vec[j_Host2]
  return term1 - term2

def FP1(in_vec, t):
  term1 = exp(-a1 * in_vec[j_Host1]) * (1 - exp(-c11 * in_vec[j_Paras1] - c21 * in_vec[j_Paras2])) * c11 * in_vec[j_Paras1] / (c11 * in_vec[j_Paras1] + c21 * in_vec[j_Paras2]) * in_vec[j_Host1]
  
  term2 = exp(-a2 * in_vec[j_Host2]) * (1 - exp(-c12 * in_vec[j_Paras1] - c22 * in_vec[j_Paras2])) * c12 * in_vec[j_Paras1] / (c12 * in_vec[j_Paras1] + c22 * in_vec[j_Paras2]) * in_vec[j_Host2]
  
  term3 = (1.0 - exp(-mP1)) * in_vec[j_Paras1]
  return term1 + term2 - term3

  
def FP2(in_vec, t):
  term1 = exp(-a1 * in_vec[j_Host1]) * (1 - exp(-c11 * in_vec[j_Paras1] - c21 * in_vec[j_Paras2])) * c21 * in_vec[j_Paras2] / (c11 * in_vec[j_Paras1] + c21 * in_vec[j_Paras2]) * in_vec[j_Host1]
  
  term2 = exp(-a2 * in_vec[j_Host2]) * (1 - exp(-c12 * in_vec[j_Paras1] - c22 * in_vec[j_Paras2])) * c22 * in_vec[j_Paras2] / (c12 * in_vec[j_Paras1] + c22 * in_vec[j_Paras2]) * in_vec[j_Host2]
  
  term3 = (1.0 - exp(-mP2)) * in_vec[j_Paras2]
  return term1 + term2 - term3

#function to calculate all changes four dimensional
def diff_4HP(in_vec, t, dim):
  temp_vec = np.zeros([dim])
  temp_vec[j_Host1] = FH1(in_vec, t)
  temp_vec[j_Host2] = FH2(in_vec, t)
  temp_vec[j_Paras1] = FP1(in_vec, t)
  temp_vec[j_Paras2] = FP2(in_vec, t)
  return temp_vec

```
##### [2.2.4] Initial setting and conditions
```{python}
dim_model2 = 4
nv_4HP0 = np.zeros([dim_model2])
nv_4HP = np.zeros([dim_model2])

nv_4HP0[j_Host1] = 1.0
nv_4HP0[j_Host2] = 1.0
nv_4HP0[j_Paras1] = 0.1
nv_4HP0[j_Paras2] = 0.1
```
#### [2.3] Solving the model
##### [2.3.2] Trial with two pairs of Host - Parasitoid system
Running the model
```{python}
t = 0  # initial condition (initial time, 0)
end_time = 2000

HP_result = np.zeros([end_time + 1, dim_model2 + 1]) #array for saving the time series generated

#Setting initial condition; it is important to start with asymmetric densities to reach quickly the chaotic attractor
nv_4HP[j_Host1] = 1.0
nv_4HP[j_Host2] = 0.5
nv_4HP[j_Paras1] = 0.1
nv_4HP[j_Paras2] = 0.2

#record initial condition
HP_result[0, 1:] = nv_4HP

#For the transient dynamics 
for i in range (1, (end_time + 1)): 
  nv_4HP = nv_4HP + diff_4HP(nv_4HP, t, dim_model2)
  t = t + 1 #update time
  HP_result[i, 0] = t
  HP_result[i, 1:] = nv_4HP

print(HP_result)
```
Plotting the time series
```{python}
plt.clf() #clear the figure setting
plt.xlim([0, end_time])
plt.ylim([0, 5])
plt.plot(HP_result[:, 0], HP_result[:, j_Host1 + 1], color = "black", label = "H1", linestyle = "solid", linewidth = 0.5)
plt.plot(HP_result[:, 0], HP_result[:, j_Host2 + 1], color = "black", label = "H2", linestyle = "dashed", linewidth = 0.5)
plt.plot(HP_result[:, 0], HP_result[:, j_Paras1 + 1], color = "red", label = "P1", linestyle = "solid", linewidth = 0.5)
plt.plot(HP_result[:, 0], HP_result[:, j_Paras2 + 1], color = "red", label = "P2", linestyle = "dashed", linewidth = 0.5)

plt.xlabel('time step t')
plt.ylabel('H1, H2, P1, P2')
plt.legend()
plt.show()

```
Converting the array to dataframe for pyEDM
```{python}
HP_result_df = pd.DataFrame(HP_result, columns = ["time", "H1", "H2", "P1", "P2"])
print(HP_result_df)
```

### [3] A simple analysis with multivariate S-map
Based on https://ushio-ecology-blog.blogspot.com/2019/12/20191225blogger0007.html

#### [3.1] Analysis with standardization
Subset of the final part of the generated time series
```{python}
HP_result_df = HP_result_df.loc[1800:1999]
print(HP_result_df)
```

##### [3.1.1] Standardization
```{python}
HP_result_mean = HP_result_df.iloc[:, 1:].mean(axis = 0, skipna = True)  # mean abundance
HP_result_sd = HP_result_df.iloc[:, 1:].std(axis = 0, skipna = True)      # SD of abundance 
print(HP_result_mean)
print(HP_result_sd)

HP_result_s = (HP_result_df.iloc[:, 1:] - HP_result_mean) / HP_result_sd # Standardized data set
#add the first column as time
HP_result_s.insert(0, "time", HP_result_df.iloc[:, 0])
print(HP_result_s)
```

##### [3.1.2] Analysis with the standardized time series data 
- Conducting univariate smap for obtaining the optimal theta

- But do not use the result, because R version determined the optimal theta by the minimum rmse while pyEDM's function returns the correlation coefficient rho only.
```{python}
result_optimal_theta_H1 = pyEDM.PredictNonlinear(dataFrame = HP_result_s, columns = "H1", target = "H1",  lib = [1, 200], pred = [1, 200], E = 4)
print(result_optimal_theta_H1)
```

Conducting multivariate Smap with the optimal theta (= 8, based on the result from rEDM)

- Note that time index for the coefficient is different from the outcome of rEDM
```{python}
multivariate_Smap_HP = pyEDM.SMap(dataFrame = HP_result_s, columns = ["H1", "H2", "P1", "P2"], target = "H1", lib = [1, 200], pred = [1, 200], theta = 8, verbose = False, showPlot = False, embedded = True)
type(multivariate_Smap_HP['coefficients'])
print(multivariate_Smap_HP['coefficients'])
```

##### [3.1.3] Plotting the results
We can see that the adjustment of the diagonal elements makes the intraspecific interaction strengths comparable to the iterspecific interaction strengths. 
```{python}
Smap_HP_coeffcients = multivariate_Smap_HP['coefficients']
delta_11 = 1 #adjustment for the diagonal elements

plt.clf() #clear the figure setting
plt.xlim([1800, 2000])
plt.ylim([-0.6, 1.6])
plt.plot(Smap_HP_coeffcients["Time"][1:], Smap_HP_coeffcients["∂H1/∂H1"][1:], color = "black", label = "H1toH1", linestyle = "solid", linewidth = 1.0, marker = "o", fillstyle = 'none')
plt.plot(Smap_HP_coeffcients["Time"][1:], Smap_HP_coeffcients["∂H1/∂H1"][1:] - delta_11, color = "#ff4b00", label = "H1toH1_adjusted", linestyle = "solid", linewidth = 1.0, marker = "s")
plt.plot(Smap_HP_coeffcients["Time"][1:], Smap_HP_coeffcients["∂H1/∂H2"][1:], color = "#4dc4ff", label = "H2toH1", linestyle = "solid", linewidth = 1.0, marker = "o")
plt.plot(Smap_HP_coeffcients["Time"][1:], Smap_HP_coeffcients["∂H1/∂P1"][1:], color = "#f6aa00", label = "P1toH1", linestyle = "solid", linewidth = 1.0, marker = "^")
plt.plot(Smap_HP_coeffcients["Time"][1:], Smap_HP_coeffcients["∂H1/∂P2"][1:], color = "#804000", label = "P2toH1", linestyle = "solid", linewidth = 1.0, marker = "D", fillstyle = 'none')


plt.xlabel('time step t')
plt.ylabel('Unadjusted and adjusted S-map coefficients')
plt.title('Figure 1: Time evolution of S-map coefficients')
plt.legend(loc = "upper center", bbox_to_anchor = (0.5, 1.0), ncol = 3)
plt.show()
```
### [5] Model02 setting
#### [5.1] Model equations for the 5-species coupled food chain ODE model
\[
\frac{dP_i}{dt} = v_i \lambda_i \frac{P_i C_i}{C_i + C_i^*} - v_i P_i, \quad (i=1,2) \tag{1}
\]

\[
\frac{dC_i}{dt} = \mu_i \kappa_i \frac{C_i R}{R + R^*} - v_i \lambda_i \frac{P_i C_i}{C_i + C_i^*} - \mu_i C_i, \quad (i=1,2) \tag{2}
\]

\[
\frac{dR}{dt} = R \left( 1 - \frac{R}{k} \right) - \sum_{i=1,2} \mu_i \kappa_i \frac{C_iR}{R + R^*}.\tag{3}
\]

#### [5.2] Parameters and settings for the 5-species coupled food chain ODE model
##### [5.2.1] Parameter value setting
```{python}
v1 = 0.1
v2 = 0.07
lambd1 = 3.2
lambd2 = 2.9
Cstar1 = 0.5
Cstar2 = 0.5
myu1 = 0.15
myu2 = 0.15
kp1 = 2.5
kp2 = 2.0
Rstar = 0.3
k = 1.2
```
##### [5.2.2] Settings for numerically solving ODEs
```{python}
j_P1 = 0
j_P2 = 1
j_C1 = 2
j_C2 = 3
j_R = 4
```
##### [5.2.3] Functions specific to the 5-sp model
```{python}
def dP1dt(in_vec, t):
  growth = v1 * lambd1 * in_vec[j_P1] * in_vec[j_C1] / (in_vec[j_C1] + Cstar1)
  mortality =  v1 * in_vec[j_P1]
  return growth - mortality 

def dP2dt(in_vec, t):
  growth = v2 * lambd2 * in_vec[j_P2] * in_vec[j_C2] / (in_vec[j_C2] + Cstar2)
  mortality = v2 * in_vec[j_P2]
  return growth - mortality 

def dC1dt(in_vec, t):
  growth = myu1 * kp1 * in_vec[j_C1] * in_vec[j_R] / (in_vec[j_R] + Rstar)
  mortality = v1 * lambd1 * in_vec[j_P1] * in_vec[j_C1] / (in_vec[j_C1] + Cstar1) + myu1 * in_vec[j_C1]
  return growth - mortality 

def dC2dt(in_vec, t):
  growth = myu2 * kp2 * in_vec[j_C2] * in_vec[j_R] / (in_vec[j_R] + Rstar)
  mortality  = v2 * lambd2 * in_vec[j_P2] * in_vec[j_C2] / (in_vec[j_C2] + Cstar2) + myu2 * in_vec[j_C2]
  return growth - mortality

def dRdt(in_vec, t):
  growth = in_vec[j_R] * (1.0 - in_vec[j_R] / k)
  mortality = myu1 * kp1 * in_vec[j_C1] * in_vec[j_R] / (in_vec[j_R] + Rstar) + myu2 * kp2 * in_vec[j_C2] * in_vec[j_R] / (in_vec[j_R] + Rstar)
  return growth - mortality

#function to calculate all coefficients, five dimensional
def diff_5sp(in_vec, t, h_interval, dim, ref_data = None):
  temp_vec = np.zeros([dim])
  temp_vec[j_P1] = dP1dt(in_vec, t)
  temp_vec[j_P2] = dP2dt(in_vec, t)
  temp_vec[j_C1] = dC1dt(in_vec, t)
  temp_vec[j_C2] = dC2dt(in_vec, t)
  temp_vec[j_R] = dRdt(in_vec, t)
  return temp_vec
```
##### [5.2.4] Initial setting and conditions
```{python}
dim_model1 = 5
nv = np.zeros([dim_model1])
deltat = 0.01
```

#### [5.3] Solving the model
##### [5.3.1] Solving the model by the RK4 method
```{python}
t = 0.0  # initial condition (initial time, 0)
end_time = 2000
tau = 5
transient_period = 1000
write_index = 1

no_step = int((end_time - transient_period) / tau) + 1
rk4_result = np.zeros([no_step, dim_model1 + 1]) #array for saving the time series generated

#Initial condition set
nv[j_R] = 1.0
nv[j_C1] = 0.5
nv[j_C2] = 0.8
nv[j_P1] = 0.7
nv[j_P2] = 0.8
print(nv)

#For the transient dynamics (RUN BURN)
for i in range(1, int(transient_period / deltat) + 1):
  rk4(in_vec = nv, out_vec = nv, time = t, h_interval = deltat, dim = dim_model1, diff_vec = diff_5sp)
  t = t + deltat # update time

#record the initial condition after BURNING phase
rk4_result[0, 0] = t
rk4_result[0, 1:] = nv

#After transient period
j = 0
for i in range(1, int((end_time - transient_period) / deltat) + 1):
  rk4(in_vec = nv, out_vec = nv, time = t, h_interval = deltat, dim = dim_model1, diff_vec = diff_5sp)
  t = t + deltat # update time
  ###record the result every "tau only
  if write_index < int(tau / deltat):
    write_index += 1
  else:
    write_index =  1
    j += 1
    rk4_result[j, 0] = t
    rk4_result[j, 1:] = nv
print(rk4_result)
```

```{python}
rk4_result_df = pd.DataFrame(rk4_result, columns = ["time", "P1", "P2", "C1", "C2", "R"])
print(rk4_result_df)
```

##### [5.3.2] Basic plot for the result
```{python}
plt.clf() #clear the figure setting
plt.xlim([transient_period, end_time])
plt.ylim([0, 2])
plt.plot(rk4_result_df["time"], rk4_result_df["C1"], color = "red", label = "C1", linestyle = "solid", linewidth = 0.5)
plt.plot(rk4_result_df["time"], rk4_result_df["C2"], color = "blue", label = "C2", linestyle = "solid", linewidth = 0.5)
plt.plot(rk4_result_df["time"], rk4_result_df["R"], color = "green", label = "R", linestyle = "solid", linewidth = 0.5)


plt.xlabel('time')
plt.ylabel('abunance')
plt.legend()
plt.show()

```
### [6] Theoretical coefficients evaluated by instantaneous Jacobian
The partial derivative of dC1/dt 


with respect to C1:
\[
\frac{\partial}{\partial C_{1}} \left( \mu_{1} K_{1} \frac{C_{1} R}{R + R^{*}} - v_{1} \lambda_{1} \frac{P_{1} C_{1}}{C_{1} + C_{1}^{*}} - \mu_{1} C_{1} \right) = K_{1} \mu_{1} \frac{R}{R + R^{*}} - v_{1} \lambda_{1} \frac{P_{1} C_{1}*}{(C_{1} + C_{1}^{*})^2} - \mu_{1}
\]

with respect to R:
\[
\frac{\partial}{\partial R} \left( \mu_{1} \kappa_{1} \frac{C_{1} R}{R + R^{*}} \right) = \kappa_{1} \mu_{1} \frac{C_{1} R^*}{(R + R^{*})^2}
\]


Function definition
```{python}
def dC1dC1(data): 
  coeff = myu1 * kp1 * data["R"] / (data["R"] + Rstar) - v1 * lambd1 * data["P1"] * Cstar1/ (data["C1"] + Cstar1) ** 2 - myu1
  return coeff 

def dC1dR(data):
  coeff = myu1 * kp1 * data["C1"] * Rstar / (data["R"] + Rstar) ** 2
  return coeff 
```


### [7] Adjusted interaction strengths with multivariate S-map
Based on https://ushio-ecology-blog.blogspot.com/2019/12/20191225blogger0007.html

#### [7.1] Analysis with standardization
##### [7.1.1] Standardization
```{python}
rk4_result_mean = rk4_result_df.iloc[:, 1:].mean(axis = 0, skipna = True)  # mean abundance
rk4_result_sd = rk4_result_df.iloc[:, 1:].std(axis = 0, skipna = True)      # SD of abundance 
print(rk4_result_mean)
print(rk4_result_sd)

rk4_result_s = (rk4_result_df.iloc[:, 1:] - rk4_result_mean) / rk4_result_sd # Standardized data set
#add the first column as time
rk4_result_s.insert(0, "time", rk4_result_df.iloc[:, 0])
print(rk4_result_s)
```

##### [7.1.2] Analaysis with the standardized time series data 

Conducting multivariate Smap with the optimal theta (= 8, based on the result from rEDM)
```{python}
multivariate_Smap_5sp = pyEDM.SMap(dataFrame = rk4_result_s, columns = ["P1", "P2", "C1", "C2", "R"], target = "C1", lib = [1, 201], pred = [1, 201], theta = 8, verbose = False, showPlot = False, embedded = True)
print(multivariate_Smap_5sp['coefficients'])
```


#### [7.2] Compare standard S-map and instantaneous interaction strengths (IIS)
```{python}
Smap_5sp_coeffcients = multivariate_Smap_5sp['coefficients']
delta_C1C1 = 1 #adjustment for the diagonal elements

plt.clf() #clear the figure setting
plt.xlim([1000, 2000])
plt.ylim([-1.6, 1.6])
plt.plot(Smap_5sp_coeffcients["Time"][1:201], Smap_5sp_coeffcients["∂C1/∂C1"][2:202] - delta_C1C1, color = "black", label = "multivariate S-map", linestyle = "solid", linewidth = 0.5)
plt.plot(Smap_5sp_coeffcients["Time"][1:201], tau*dC1dC1(rk4_result_df[1:201]), color = "blue", label = "IIS", linestyle = "solid", linewidth = 0.5)

plt.xlabel('time')
plt.ylabel('S-map coefficients: C1 -> C1')
#plt.title('')
plt.legend(loc = "upper center", bbox_to_anchor = (0.5, 1.0))
plt.show()
```
### [8] Deriving mathematically correct interaction coefficients

#### [8.2] Solving the model
##### [8.2.1] Preparing the fine solution of the ODEs without perturbation
Since rk4() includes the evaluation of derivatives at t + deltat/2, we need to have numerical solutions with the high resolution with deltat/2
```{python}
t = 0.0  # initial condition (initial time, 0)
end_time = 2000
tau = 5
transient_period = 1000
write_index = 1
deltat_fine = deltat * 0.5

no_step_fine = int((end_time - transient_period) / deltat_fine) + 1
rk4_result_fine = np.zeros([no_step_fine, dim_model1 + 1]) #array for saving the time series generated

#Initial condition set
nv_rk4_fine = np.zeros([dim_model1])

#Initial condition set
nv_rk4_fine[j_R] = 1.0
nv_rk4_fine[j_C1] = 0.5
nv_rk4_fine[j_C2] = 0.8
nv_rk4_fine[j_P1] = 0.7
nv_rk4_fine[j_P2] = 0.8

#For the transient dynamics (RUN BURN)
for i in range(1, int(transient_period / deltat_fine) + 1):
  rk4(in_vec = nv_rk4_fine, out_vec = nv_rk4_fine, time = t, h_interval = deltat_fine, dim = dim_model1, diff_vec = diff_5sp)
  t = t + deltat_fine # update time

#record the initial condition after BURNING phase
rk4_result_fine[0, 0] = t
rk4_result_fine[0, 1:] = nv_rk4_fine

#After transient period, recording every time step
j = 0
for i in range(1, int((end_time - transient_period) / deltat_fine) + 1):
  rk4(in_vec = nv_rk4_fine, out_vec = nv_rk4_fine, time = t, h_interval = deltat_fine, dim = dim_model1, diff_vec = diff_5sp)
  t = t + deltat_fine # update time
  j += 1
  rk4_result_fine[j, 0] = t
  rk4_result_fine[j, 1:] = nv_rk4_fine

print(rk4_result_fine)
```

```{python}
rk4_result_fine_df = pd.DataFrame(rk4_result_fine, columns = ["time", "P1", "P2", "C1", "C2", "R"])
print(rk4_result_fine_df)
```

Comparing the solutions 
```{python}

plt.clf() #clear the figure setting
plt.xlim([1000, 2000])
plt.ylim([-0.1, 1.6])
plt.plot(rk4_result_df["time"], rk4_result_df["C1"], color = "red", label = "default delta-t", linestyle = "solid", linewidth = 0.5)
plt.plot(rk4_result_fine_df["time"], rk4_result_fine_df["C1"], color = "black", label = "half delta-t", linestyle = "solid", linewidth = 0.5)

plt.xlabel('time')
plt.ylabel('C1 abundance')
#plt.title('')
plt.legend(loc = "upper center", bbox_to_anchor = (0.5, 1.0))
plt.show()
```

##### [8.2.3] Functions to calculate the interaction strength
```{python}
#ref_data: the dataset that includes the fine resolution solution of the target nonlinear ODE systems, of which record started at absolute time = transient_period
#deltax: the size of perturbation at t = h_interval/2
#start_t: usually specified as transient_period
#tau: the interval for the calculation
#dim: dimension of ODE system
#diff_vec: the vector field generated by the linearized ODEs with non-perturbed solution or the original nonlinear ODE

#This is a function that directly used the nonlinear ODE with the perturbed initial conditions
def direct_coeff(ref_data, deltax, start_t, tau, h_interval, dim, diff_vec):
  #matrix to stock results
  nv_nonlinear_p = np.zeros([dim, dim])
  
  tzero = start_t - transient_period + h_interval / 2  # initial condition (initial time, t_abs = start_t)
  t_index_start = int((tzero / h_interval) * 2) - 1
  t_index_end = t_index_start + int((tau / h_interval) * 2)
  
  #Initial condition 
  nv_nonlinear_p[:, j_P1] = ref_data.iloc[t_index_start, 1:] + [deltax, 0, 0, 0, 0]
  nv_nonlinear_p[:, j_P2] = ref_data.iloc[t_index_start, 1:] + [0, deltax, 0, 0, 0]
  nv_nonlinear_p[:, j_C1] = ref_data.iloc[t_index_start, 1:] + [0, 0, deltax, 0, 0]
  nv_nonlinear_p[:, j_C2] = ref_data.iloc[t_index_start, 1:] + [0, 0, 0, deltax, 0]
  nv_nonlinear_p[:, j_R] = ref_data.iloc[t_index_start, 1:] + [0, 0, 0, 0, deltax]    
 
  end_time = tau
  
  for j in range(0, dim): 
    t = tzero
    for i in range(1, int(end_time / deltat) + 1):
      rk4(nv_nonlinear_p[:, j], nv_nonlinear_p[:, j], t, h_interval, dim, diff_vec)
      t = t + h_interval # update time (starting from the middle)
  
  #Calculate the difference between the perturbed flow and unperturbed flow
  for j in range(0, dim):
    nv_nonlinear_p[:, j] = nv_nonlinear_p[:, j] - ref_data.iloc[t_index_end, 1:]  
  
  return nv_nonlinear_p / deltax

```


##### [8.2.5] Nonlinear ODE model for calculating all points from 1000 to 2000 with interval 
The blokc using mclapply() in R is simply replaced by for loop in python, since the computational time is not large even in a laptop Windows PC.
```{r eval=FALSE}
direct_coeff_5sp <- list()

start_time <- vector()
transient_period <- 1000

direct_coeff_5sp <- mclapply(1:200, function(i) {
  start_time[i] <- transient_period + (i - 1)*tau
  result_nonlinear <- direct_coeff(rk4_result_fine, deltax = 0.0001, start_t = start_time[i], tau, deltat, dim = 5, diff_5sp)
  result_nonlinear
}, mc.cores = 1)

direct_coeff_5sp[[5]]
```
Python codes and results
```{python}
no_step = 200
direct_coeff_5sp = np.zeros([no_step, dim_model1, dim_model1])
for i in range(0, no_step):
  start_t_i = transient_period + i * tau
  direct_coeff_5sp[i, :, :] = direct_coeff(rk4_result_fine_df, deltax = 0.0001, start_t = start_t_i, tau = 5, h_interval = deltat, dim = dim_model1, diff_vec = diff_5sp)
  
print(direct_coeff_5sp[4, :, :])
```

### [10] Comparisons for 5sp model

#### [10.1] Compare standard S-map, IIS, and CIS by direct evaluation (but without MDR S-map)
CIS by Direct evaluation: direct_IS (Eqn.19)
```{python}
#normalization by SD
sd_adj_C1R = rk4_result_sd[j_R] / rk4_result_sd[j_C1]
delta_C1C1 = 1

direct_dC1dC1 = direct_coeff_5sp[:, j_C1, j_C1] - delta_C1C1
direct_dC1dR = sd_adj_C1R * direct_coeff_5sp[:, j_C1, j_R]
```
##### [10.1.2] Comparison for the diagonal element (Fig. 3a: C1 -> C1): corresponding to ggplot in R
```{python}
plt.clf() #clear the figure setting
plt.xlim([1400, 1800])
plt.ylim([-3, 2.0])
plt.plot(Smap_5sp_coeffcients["Time"][80:161], Smap_5sp_coeffcients["∂C1/∂C1"][81:162] - delta_C1C1, color = "#ff4b00", label = "01std_smapC1toC1", linestyle = "solid", linewidth = 0.5, marker = "s")
plt.plot(Smap_5sp_coeffcients["Time"][80:161], direct_dC1dC1[80:161], color = "#f6aa00", label = "03CIS_dC1dC1", linestyle = "solid", linewidth = 0.5, marker = "^")
plt.plot(Smap_5sp_coeffcients["Time"][80:161], tau*dC1dC1(rk4_result_df[80:161]), color = "#804000", label = "04IIS_dC1dC1", linestyle = "solid", linewidth = 0.5, marker = "D", fillstyle = 'none')

plt.xlabel('time')
plt.ylabel('S-map coefficients: C1 -> C1')
plt.title('Time evolution: C1 -> C1')
plt.legend(loc = "upper center", bbox_to_anchor = (0.5, 1.0), ncol = 3)
plt.show()
```

##### [10.1.3] Comparison for the diagonal element (Fig. 3c:R -> C1): corresponding to ggplot in R
```{python}
plt.clf() #clear the figure setting
plt.xlim([1400, 1800])
plt.ylim([0, 9])
plt.plot(Smap_5sp_coeffcients["Time"][80:161], Smap_5sp_coeffcients["∂C1/∂R"][81:162] - delta_C1C1, color = "#ff4b00", label = "01std_smapRtoC1", linestyle = "solid", linewidth = 0.5, marker = "s")
plt.plot(Smap_5sp_coeffcients["Time"][80:161], direct_dC1dR[80:161], color = "#f6aa00", label = "03CIS_dC1dR", linestyle = "solid", linewidth = 0.5, marker = "^")
plt.plot(Smap_5sp_coeffcients["Time"][80:161], tau*dC1dR(rk4_result_df[80:161]), color = "#804000", label = "04IIS_dC1dR", linestyle = "solid", linewidth = 0.5, marker = "D", fillstyle = 'none')

plt.xlabel('time')
plt.ylabel('S-map coefficients: R -> C1')
plt.title('Time evolution: R -> C1')
plt.legend(loc = "upper center", bbox_to_anchor = (0.5, 1.0), ncol = 3)
plt.show()
```

